---
title: "Final Project"
author: "Travis Froberg"
date: "10/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown: Collinearity and Variable Selection

```{r}
library(car)
library(nortest)
library('olsrr')
library(MASS)
library(plyr)
library(leaps)
library(tweedie)
library(statmod)
library(aTSA)
```

```{r}
setwd("C:\\Users\\malin\\Documents\\GLM Class - MTU")
LoanData <- read.csv("IBRD_Loan_Data.csv")
attach(LoanData)
LoanData <- data.frame(
                       LoanData$Disbursed.Amount,
                       LoanData$Sold.3rd.Party,
                       LoanData$Region,
                       LoanData$Loan.Status,
                       Original.Principal.Amount,
                       Borrower.s.Obligation,
                       Cancelled.Amount,
                       LoanData$Interest.Rate
)
LoanData <- na.omit(LoanData)
# str(LoanData)
colnames(LoanData) <- c("Disbursed_Amount", "Sold_3rd_Party", "Region",
                        "Loan_Status", "Original_Principal_Amount",
                        "Borrowers_Obligation", "Cancelled_Amount", "Interest_Rate")
# attach(LoanData)
LoanData$Disbursed_Amount <- LoanData$Disbursed_Amount/1000000
LoanData$Sold_3rd_Party <- LoanData$Sold_3rd_Party/1000000
LoanData$Original_Principal_Amount <- LoanData$Original_Principal_Amount/1000000
LoanData$Borrowers_Obligation <- LoanData$Borrowers_Obligation/1000000
LoanData$Cancelled_Amount <- LoanData$Cancelled_Amount/1000000
LoanData$Interest_Rate <- LoanData$Interest_Rate/1000000
LoanData$Region <- as.factor(LoanData$Region)
LoanData$Loan_Status <- as.factor(LoanData$Loan_Status)

# levels(LoanData2$Region)
levels(LoanData2$Loan_Status)
LoanData$Region <- recode(LoanData$Region, "c('AFRICA EAST', 'AFRICA WEST') = 'East_and_West_Africa'")
LoanData$Region <- recode(LoanData$Region, "c('EUROPE AND CENTRAL ASIA', 'MIDDLE EAST AND NORTH AFRICA', 'SOUTH ASIA') = 'EUROPE, MIDDLE EAST, NORTH AFRICA, AND CENTRAL AND SOUTH ASIA'")
# levels(X3)
### LoanData2 <- subset(LoanData, LoanData$Region!="SOUTH ASIA" & LoanData$Region!="MIDDLE EAST AND NORTH AFRICA" & LoanData$Region!="OTHER" & LoanData$Region!="EAST ASIA AND PACIFIC" )
# View(LoanData2)
LoanData2 <- LoanData
LoanData2$Region <- as.factor(LoanData2$Region)
# levels(LoanData2$X3)

# attach(LoanData2)
LoanData2$Loan_Status <- as.factor(LoanData2$Loan_Status)

LoanData2$Loan_Status <- recode(LoanData2$Loan_Status, "c('Approved', 'Disbursing', 'Disbursing&Repaying', 'Effective', 'Fully Disbursed', 'Fully Transferred', 'Repaying', 'Signed') = 'Active_(not_fully_repaid_or_cancelled)'")
LoanData2$Loan_Status <- recode(LoanData2$Loan_Status, "c('Fully Cancelled', 'Terminated') = 'Cancelled'")

LoanData2$Region <- relevel(LoanData2$Region, ref = 'EUROPE, MIDDLE EAST, NORTH AFRICA, AND CENTRAL AND SOUTH ASIA')

LoanData2$Loan_Status <- relevel(LoanData2$Loan_Status, ref = 'Cancelled')

LoanData2$Region <- factor(LoanData2$Region)
LoanData2$Region <- (LoanData2$Region)
LoanData2$Loan_Status <- factor(LoanData2$Loan_Status)
levels(LoanData2$Region)
# View(LoanData2)
table(LoanData2$Region)
```
```{r}
plot(LoanData2$Interest_Rate)
range(LoanData2$Interest_Rate)
densityPlot(LoanData2$Interest_Rate)
mean(LoanData2$Interest_Rate)
mean(LoanData2$Interest_Rate)^2
mean(LoanData2$Interest_Rate)^3
var(LoanData2$Interest_Rate)

# Best so Far 
mean(LoanData2$Interest_Rate)^1.1
mean(LoanData2$Interest_Rate)^1.268137
mean(LoanData2$Interest_Rate)*(1-mean(LoanData2$Interest_Rate))
```

```{r}
scatterplotMatrix(~ Interest_Rate+ Disbursed_Amount + Cancelled_Amount + Original_Principal_Amount, LoanData2, smooth = FALSE )
boxplot(LoanData3$Interest_Rate~LoanData2$Region)
boxplot(LoanData3$Interest_Rate~LoanData2$Loan_Status, xlab = "Loan Status", ylab = "Interest Rate")
plot(LoanData2$Interest_Rate, ylab = "Interest Rate")
```
# Summary of Possible Collinearity 
As can be seen in the scatterplot above, X1 and X5, X1 and X6, and X5 and X6 all showed an obvious linear trend in their relations. This is expected because all three of these variables are measuring similar things. X1 is the amount of the loan that has been disbursed to a particular country, X5 is the principal amount of the original loan, and X6 is the amount the borrowing company still owes. 

It makes sense that a higher original loan amount (X5) corresponds to a higher disbursed amount (X1) of the loan. And a higher disbursed amount (X1) would correspond with a higher amount still owed (X6). 

# Estimation of Index Parameter

```{r}
out <- tweedie.profile(Interest_Rate ~ Region, do.plot=TRUE, data =LoanData2, xi.vec = seq(1.02, 1.4, length = 19))
out$xi.max
```

```{r}
# profile likelihood 
xi.est <- out$xi.max # estimated index parameter
out$ci # ci of index parameter
out$phi.max # dispersion parameter
```

# GLM Models 
```{r}
mod1 <- glm(Interest_Rate ~ Disbursed_Amount + 
                       # Original_Principal_Amount + 
                       # Borrowers_Obligation +
                        Cancelled_Amount +
                       # Sold_3rd_Party + 
                        Region +
                        Loan_Status,
                        data = LoanData2, 
   family = tweedie(var.power = xi.est , link.power = 0))
printCoefmat(coef(summary(mod1)))
summary(mod1)
# adf.test(mod1)
# levels(LoanData2$Region)
```
```{r}
# Model with the variable Borrowers Obligation. 
mod1.1 <- glm(Interest_Rate ~ Disbursed_Amount + 
                         Original_Principal_Amount + 
                       # Borrowers_Obligation +
                        Cancelled_Amount +
                       # Sold_3rd_Party + 
                        Region +
                        Loan_Status,
                        data = LoanData2, 
   family = tweedie(var.power = xi.est , link.power = 0))
printCoefmat(coef(summary(mod1.1)))
summary(mod1.1)
```
```{r}
# perform the F-test
dev2 <- deviance(mod1) - deviance(mod1.1)
df1 <- df.residual(mod1) - df.residual(mod1.1)
df2 <- df.residual(mod1.1)
f <- dev2 / df1 / summary(mod1.1)$dispersion
p.value <- 1 - pf(f, df1 = df1, df2 = df2)
c(f = f, p = p.value, df1 = df1, df2 = df2) 
```

```{r}
# Model without the variable Cancelled Amount 
mod1.2 <- glm(Interest_Rate ~ Disbursed_Amount + 
                         Original_Principal_Amount + 
                       # Borrowers_Obligation +
                       # Cancelled_Amount +
                       # Sold_3rd_Party + 
                        Region +
                        Loan_Status,
                        data = LoanData2, 
   family = tweedie(var.power = xi.est , link.power = 0))
printCoefmat(coef(summary(mod1.2)))
summary(mod1.2)
```
```{r}
# perform the F-test
dev2 <- deviance(mod1.2) - deviance(mod1.1)
df1 <- df.residual(mod1.2) - df.residual(mod1.1)
df2 <- df.residual(mod1.1)
f <- dev2 / df1 / summary(mod1.1)$dispersion
p.value <- 1 - pf(f, df1 = df1, df2 = df2)
c(f = f, p = p.value, df1 = df1, df2 = df2) 
```
```{r}
# Model without Cancelled Amount and with Original Principal Amount 
mod1.3 <- glm(Interest_Rate ~ Disbursed_Amount + 
                         Original_Principal_Amount + 
                       # Borrowers_Obligation +
                       # Cancelled_Amount +
                       # Sold_3rd_Party + 
                        Region +
                        Loan_Status,
                        data = LoanData2, 
   family = tweedie(var.power = xi.est , link.power = 0))
printCoefmat(coef(summary(mod1.3)))
summary(mod1.3)
# adf.test(mod1)
# levels(LoanData2$Region)
```
```{r}
# diagnostic plots 
rQ <- qresid(mod1.3) # quantile residual
mu <- fitted(mod1.3) # fitted values
lp <- predict(mod1.3, type = "link") # linear predictor
z <- resid(mod1.3, type = "working") + lp # working response
rP <- resid(mod1.3, type = "partial") # partial residuals
# (3): plot of quantile residuals against transformed mu
par(mfrow = c(2, 3))
scatter.smooth(rQ ~ mu^(2-xi.est)/2, las = 1, ylab = "Quantile Residuals", 
   xlab = "Fitted Values (Transformed)")
  # main = "A: Quantile Residuals vs Fitted Values")
# (3): plot of working response against linear predictor
scatter.smooth(z ~ lp, las = 1, ylab = "Working Responses", 
   xlab = "Linear Predictors") 
   # main = "C: Working Responses vs Linear Predictors")
# (3): Q-Q plot of quantile residuals
qqnorm(rQ, las = 1, main = "E: Q-Q plot")
qqline(rQ)
# (3): outliers and influential observations
sort(rQ)
cD <- cooks.distance(mod1.3)
plot(sort(cD), ylab = "Cook's Distance")

# 
# plot(cD ~ turbines$Hours, las = 1, 
#    ylab = "Cook's Distance", xlab = "Hours", 
#    main = "F: Cook's Distance\n vs Covatiate Hours")
# maxrq <- which.max(abs(rQ))
# maxd <- which.max(cD)
# points( x = turbines$Hours[maxrq], y = cD[maxrq], col = 3, pch = 16) 
# points( x = turbines$Hours[maxd], y = cD[maxd], col = 4, pch = 16)

```

```{r}
mod2.loan <- glm(Interest_Rate ~ 
                        Borrowers_Obligation +
                        Sold_3rd_Party + 
                        Region +
                        Loan_Status +
                         + Cancelled_Amount, data = LoanData2, 
                         family = inverse.gaussian)# (link = "log") )
printCoefmat(coef(summary(mod2.loan)))
```

```{r}
mod3.loan <- glm(Interest_Rate ~ Disbursed_Amount +
                        # Original_Principal_Amount * 
                       #  Borrowers_Obligation +
                       #  Sold_3rd_Party + 
                        Region +
                        Loan_Status +
                         + Cancelled_Amount, data = LoanData2, 
   family = binomial(link = "logit"))
summary(mod3.loan)
```

```{r}
mod4.loan <- glm(Interest_Rate ~ Disbursed_Amount * Original_Principal_Amount * 
                        Borrowers_Obligation +
                        Sold_3rd_Party + 
                        Region +
                        Loan_Status +
                         + Cancelled_Amount, data = LoanData2, 
   family = poisson(link = "log"))
summary(mod4.loan)
```

